An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-03-30 EconGrowthSeminar.Rnw
Several small changes and edits in preparation for the seminar.

% This is a sequence of two slides with 3-D graphs. 
% The first just shows the surface. The second also shows the optimum point in the surface.
\begin{frame}{$SSE$ Minimization}{US single-factor with $Q$} %---------------------------- 
<<3D_Graph_US_SF_Q, fig.width=presentationGraph1ColWidth*1.3, fig.height=presentationGraph1RowHeight*1.3, eval=TRUE>>=
sf3DSSEGraph("US", "Q", showOpt=FALSE)
@
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putFlagUpperRight{US}{0.055}  
\end{frame}

\begin{frame}{$SSE$ Minimization}{US single-factor with $Q$} %---------------------------- 
<<3D_Graph_US_SF_Q_With_Opt_Point, fig.width=presentationGraph1ColWidth*1.3, fig.height=presentationGraph1RowHeight*1.3, eval=TRUE>>=
sf3DSSEGraph("US", "Q", showOpt=TRUE)
@
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putFlagUpperRight{US}{0.055}  
\end{frame}

@@

    \item{$CI_{\lambda}$ smaller than $CI_{\gamma}$ for all countries}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Tentative Conclusions} %----------------------------
  \begin{itemize}
  \item{Including energy in Cobb-Douglas and CES production functions is unwarranted \\
        \scriptsize
        Preliminary statistical evidence indicates that energy parameters are not significant for most countries \\Why? Energy is \emph{always already} embedded in capital stock ($K$) and labor ($L$).
        \normalsize}
  \item{Including energy in LINEX production function is required \\
        \scriptsize
        The foundation of LINEX is an energy term.\\Statistical evidence indicates that LINEX parameters $a_0$ and $c_t$ are significant for most countries.\\However, LINEX fits are worse overall than Cobb-Douglas or CES.
        \normalsize}
  \end{itemize}
\end{frame}

\begin{frame}{The Question} %----------------------------
  \begin{center}
    Which economic growth model best describes (statistically) the role of energy in economic growth for different types of economies?
  \end{center}
\end{frame}

@@

    \pause
    Answer: Cobb-Douglas or CES without energy
  \end{center}
\end{frame}

\begin{frame}{Extra Credit?} %----------------------------
  \begin{center} Reflect dynamics of energy cost share and price spikes \end{center}
    \vspace{\baselineskip}

    \pause
  \begin{itemize}
    \item{Price, cost share exogenous to all models}
    \pause
    \item{Models (CD and CES without energy) could be \emph{included within} a model that endogenizes price and cost share}
    \pause
    \item{Still searching for a modeling approach that captures price and cost share dynamics \\ \scriptsize System dynamics modeling? \normalsize}
    \pause
    \item{No extra credit!}
  \end{itemize}
\end{frame}

\begin{frame}{Next Steps (Summer 2013)} %----------------------------
  \begin{itemize}
    \item{Estimate statistical properties for \emph{all} fitting parameters}
    \vspace{\baselineskip}
    \item{Incorporate all constraints for Cobb-Douglas model}
    \vspace{\baselineskip}
    \item{Complete paper for \emph{Energy Economics}}
  \end{itemize}  
\end{frame}

>>>>

% This is a sequence of two slides with 3-D graphs. 
% The first just shows the surface. The second also shows the optimum point in the surface.
\begin{frame}{$SSE$ Minimization}{US single-factor with $Q$} %---------------------------- 
<<3D_Graph_US_SF_Q, fig.width=presentationGraph1ColWidth*1.3, fig.height=presentationGraph1RowHeight*1.3>>=
sf3DSSEGraph("US", "Q", showOpt=FALSE)
@
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putFlagUpperRight{US}{0.055}  
\end{frame}

\begin{frame}{$SSE$ Minimization}{US single-factor with $Q$} %---------------------------- 
<<3D_Graph_US_SF_Q_With_Opt_Point, fig.width=presentationGraph1ColWidth*1.3, fig.height=presentationGraph1RowHeight*1.3>>=
sf3DSSEGraph("US", "Q", showOpt=TRUE)
@
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putFlagUpperRight{US}{0.055}  
\end{frame}

@@

    \item{$CI_{\lambda}$ smaller than $CI_{\gamma}$ for all countries\footnote{\tiny as percentage of $0.0 < \lambda < 0.1$ and $0 < \gamma < 1$ \normalsize}}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Summary of Analysis} %----------------------------
  \begin{itemize}
  \item{Including energy ($E$) in Cobb-Douglas and CES production functions appears to be unwarranted}
  \pause
    \begin{itemize}
      \scriptsize
      \item{Energy parameters less significant than others for most countries}
      \pause
      \item{Why? Energy is \emph{always already} embedded in capital stock ($K$) and labor ($L$)}
      \normalsize
    \end{itemize}
  \vspace{\baselineskip}
  \pause
  \item{Including energy ($E$) in LINEX production function is required}
  \pause
    \begin{itemize}
      \scriptsize
      \item{One factor of production: energy}
      \pause
      \item{Statistical evidence indicates that LINEX parameters $a_0$ and $c_t$ are significant for most countries}
      \pause
      \item{Unfortunately, LINEX fits are worse overall than Cobb-Douglas or CES}
      \normalsize
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{The Question} %----------------------------
  \begin{center}
    Which economic growth model best describes (statistically) the role of energy in economic growth for different types of economies?
  \end{center}
  \vspace{\baselineskip}

\end{frame}

@@

    Answer: Cobb-Douglas or CES without energy
  \end{center}
\end{frame}

\begin{frame}{Extra Credit?} %----------------------------
  \begin{center} Reflect dynamics of energy cost share and price spikes? \end{center}
    \vspace{\baselineskip}

    \pause
  \begin{itemize}
    \item{Price, cost share exogenous to all models}
    \pause
    \item{No extra credit!}
    \pause
    \vspace{\baselineskip}

    \item{Models (CD and CES without energy) could be \emph{included within} a model that endogenizes price and cost share}
    \pause
    \item{Still searching for a modeling approach that captures price and cost share dynamics \\ \scriptsize System dynamics modeling? \normalsize}
  \end{itemize}
\end{frame}

\begin{frame}{Next Steps (Summer 2013)} %----------------------------
  \begin{itemize}
    \item{Verify and improve estimates of statistical properties for \emph{all} fitting parameters}
    \vspace{\baselineskip}
    \item{Incorporate all constraints for Cobb-Douglas model}
    \vspace{\baselineskip}
    \item{Complete paper for \emph{Energy Economics}}
  \end{itemize}  
\end{frame}
